\chapter{Micro-architectural Event Tracking}

Version 1.9.1 of the RISC-V Privileged Architecture adds support for Hardware Performance Monitor (HPM) counters.\footnote{Future efforts may add some counters into a memory-mapped access region. This will open up the ability to track events that, for example, may not be tied to any particular core (like last-level cache misses).}  The HPM support allows a nearly infinite number of micro-architectural events to be multiplexed onto up to 29 physical counters.  The privilege access levels can also be set on a per-counter basis.

\begin{table}[htp]
\caption{Uarch Events}
\begin{center}
\begin{tabular}{|c|c|}
\hline
Number & Event \\
\hline
\hline
1 & Committed Branches  \\
\hline
2 & Committed Branch Mispredictions\\
\hline
... & and many more (see core.scala) \\
\hline
\end{tabular}
\end{center}
\label{table:uarchcounters}
\end{table}%

The available events can be modified in {\tt core.scala} as desired.\footnote{Note: {\tt io.events(0)} will get mapped to being {\tt Event \#1} (etc.) from the point of view of the RISC-V software.}  It is then up to machine-level software to set the privilege access level and the {\em event selectors} for each counter as desired.

\begin{center}
\begin{minipage}{0.66\textwidth}
\begin{lstlisting}[caption=Setting the privilege level and event selectors for the HPM counters.]
static void mstatus_init()
{
   // Enable user/supervisor use of perf counters
   write_csr(mucounteren, -1);
   write_csr(mscounteren, -1);

   // set events 1 and 2 to counters 3 and 4.
   write_csr(mhpmevent3, 1);
   write_csr(mhpmevent4, 2);
\end{lstlisting}\label{ref:code_hpm_init}
\end{minipage}
\end{center}


\section{Reading HPM Counters in Software}

The Code Example \ref{ref:code_uarch} demonstrates how to read the value of any CSR register from software.
  
  
\begin{center}
\begin{minipage}{0.66\textwidth}
\begin{lstlisting}[caption=Reading a CSR register.]
#define read_csr_safe(reg) ({ register long __tmp asm("a0"); \   
  asm volatile ("csrr %0, " #reg : "=r"(__tmp)); \               
  __tmp; })             
  
  long csr_cycle  = read_csr_safe(cycle);
  long csr_instr  = read_csr_safe(instret);
  long csr_hpmc3  = read_csr_safe(hpmcounter3);
  ...
  long csr_hpmc31 = read_csr_safe(hpmcounter31);
  
\end{lstlisting}\label{ref:code_uarch}
\end{minipage}
\end{center}


